Project Protocol

### 2021-01-11 paper/paper.Rmd

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used, the sizes of areas can lead to the misinterpretation of the distribution. The visualization method used to present geospatial data will influence the understanding of the distribution derived by map users. Choosing an effective alternative display could positively influence the communication of the spatial distribution. The hexagon tile map is presented as an alternative display for visualizing population related distributions effectively. Visual inference is used to measure the power of design, and the choropleth is used as a comparison. The hexagon tile map display is also tested using a distribution that is directly related to the geography, with values monotonically decreasing from the North-West to South-East areas of Australia. This study finds the single map in a hexagon tile map lineup that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
dev: pdf
bibliography: paper.bib
output:
  bookdown::pdf_book: 
      base_format: rticles::ieee_article
      number_sections: true
editor_options: 
  chunk_output_type: inline
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

>>>>

---
title: "Comparing the Effectiveness of the Choropleth Map with a Hexagon Tile Map for Communicating Cancer Statistics"
affiliation:
  author-columnar: true         ## one column per author
  #institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields
  institution:
    - name: Queensland University of Technology
      department: Science and Engineering Faculty
      location: Brisbane, Australia
      email: stephanie.kobakian@qut.edu.au
      mark: 1
      author:
        - name: Stephanie Kobakian
    - name: Monash University
      department: Econometrics and Business Statistics Faculty
      location: Melbourne, Australia
      email: dicook@monash.edu
      mark: 1
      author:
        - name: Dianne Cook
keywords: ["statistics", "visual inference", "geospatial", "population"]
abstract: |
  The choropleth map display is commonly used for communicating spatial distributions across geographic areas. However, when choropleths are used the size of the geographic units will influence the understanding of the distribution derived by map users. The hexagon tile map is presented as an alternative display for visualizing population related distributions effectively. Visual inference is used to measure the power of the hexagon tile map design, and the choropleth is used as a comparison. The hexagon tile map display is tested using a distribution that is directly related to the geography, with values monotonically increasing from the North-West to South-East areas of Australia. This study finds in a hexagon tile map lineup the single map that contains a population related distribution is detected with greater probability than the same data displayed in a choropleth map. These findings should encourage map creators to implement alternative displays and consider a hexagon tile map when presenting spatial distributions of heterogeneous areas.
dev: pdf
bibliography: paper.bib
output:
  bookdown::pdf_book: 
      base_format: rticles::ieee_article
      number_sections: true
editor_options: 
  chunk_output_type: inline
link-citations: true
citation_package: biblatex
csl: springer-vancouver.csl
header-includes: \usepackage{graphicx}
---

### 2021-01-11 paper/paper.tex

Participants provided their level of certainty regarding their choice using a five point scale.
Unlike the accuracy and speed of responses that were derived during the data processing phase, this was a subjective
assessment by the participant prompted by the question: `How certain are you about your choice?'.
Figure. \ref{certainty} shows the amount of times participants provided each level of certainty. This was separated for each combination of trend models and display type, and colored depending on whether a participant correctly detected the data plot in the lineup.
Participants chose 4 or 5 often when viewing the population related trends in the choropelth display, even though they were often incorrect when viewing an All Cities trend and overwhelmingly incorrect for the Three Cities trend. This shows overconfidence in their detection ability when using a choropleth map display. Participants were less likely to be certain when their choice was incorrect and they were viewing a hexagon tile map.
For each trend model, participants were more likely to doubt their choice and choose 1 or 2 in the hexagon tile map displays, even though many had made the correct choice.

@@

\leavevmode\hypertarget{ref-atlas}{}%
2. Queensland CC Australian Cancer Atlas (https://atlas.cancer.org.au).

>>>>

Participants provided their level of certainty regarding their choice using a five point scale.
Unlike the accuracy and speed of responses that were derived during the data processing phase, this was a subjective
assessment by the participant prompted by the question: `How certain are you about your choice?'.
Fig. \ref{fig:certainty} shows the amount of times participants provided each level of certainty. This was separated for each combination of trend models and display type, and colored depending on whether a participant correctly detected the data plot in the lineup.
Participants often chose 4 or 5 when viewing the population related trends in the choropelth display, even though they were often incorrect when viewing an All Cities trend and overwhelmingly incorrect for the Three Cities trend. This shows overconfidence in their detection ability when using a choropleth map display. Participants were less likely to be certain when their choice was incorrect and they were viewing a hexagon tile map.
For each trend model, participants were more likely to doubt their choice and choose 1 or 2 in the hexagon tile map displays, even though many had made the correct choice.

@@

\leavevmode\hypertarget{ref-atlas}{}%
2. Cancer Council Queensland Australian Cancer Atlas (https://atlas.cancer.org.au).

### 2021-01-11 protocol.Rmd

Factor combinations to be examined by each participant amount to 6 (2x3) lineup displays. A participant cannot see the same data for both plot types. Four simulated sets of data will be generated for each treatment. This will generate 24 lineups (12 will be geographic maps, and 12 will be hexagon tile maps). Participants will evaluate 12 lineups, 6 of each plot type.

@@

##	Recruitment strategies, timeframe 
Recruitment will take place via advertising on the Figure-Eight dashboards.
The Figure-Eight platform will advertise this as a "job" to potential participants, they will not be under any obligation to complete this task. We keep advertising until enough participants have completed the survey, then remove the advertisements and close the survey.

@@

##	Participant withdrawal
Participants can withdraw from the experiment at any time by quitting the survey.

@@

Test questions are the most important quality control mechanism in the Figure-Eight Platform. Participants will complete them to practice and prepare for the experiment questions, they will need to reach an accuracy threshold we will set, to be allowed to participate in answering the experiment survey questions.

The Figure Eight procedures are followed: *Contributors that fail quiz mode are not paid and are permanently disqualified from working on your job. Participants will need to achieve a pre-determined accuracy threshold to be allowed to answer the questions in our survey.*

@@

Data will be collected through survey distribution via a web application.
Participants will access the application via a link from the Figure-Eight "job".

@@

```{r echo=FALSE, eval=TRUE, fig.height=5, fig.width=6, out.width="70%"}
# Code to compute the power for given sample size and alternative p
library(tidyverse)
library(nullabor)
library(viridis)
library(knitr)
n <- 20:50
pA <- seq(1/12, 1/3, 0.002)
pow <- sample_size(n, 12, pA)
#g <- expand.grid(n, pA)
#pow <- tibble(n=g[,1], k=as.integer(ceiling(g[,1]/12)), pA=g[,2]) 
#pow <- pow %>% 
#  mutate(prob = pbinom(k-1, n, pA, lower.tail=FALSE)) 
ggplot(pow, aes(x=n, y=pA, fill=prob, group=pA)) + 
  geom_tile() +
  scale_fill_viridis_c("power") +
  geom_hline(yintercept=0.1666667, colour="black") +
  ylab("detect rate (pA)") + xlab("sample size (n)") +
  theme_bw()
pow %>% filter(near(pA, 0.167, 0.001)) %>% filter(prob>0.7) %>% top_n(n=10) %>% kable(digits=3)
```

@@

Research expectations today require that publications use reproducible research protocols, requiring data to be available. De-identified, and privatised data will be made available will be made available via QUT's Research Data Finder, and, if required by the publisher, once the results are published also available on the journal article web site.

@@

Datasets produced from the project and that will *not* be available for public access will contain original demographic information about the participants and only the anonymous Figure-Eight contributor ID.

Datasets produced from the project and that *will be* available for public access on github will be de-identified and only contain the responses to survey.
If there are participants with unique sets of demographics, their data will be combined with larger categories to minimise the risks of reidentification.

@@

@@

To ensure records are retained during and following the study the responses will be stored on the google drive immediately following the submission by participants. At the conclusion of the survey, the data will remain in the password protected google drive account.

During the analysis phase the researchers Stephanie and Dianne will make copies of the survey responses to their work devices (laptop computers), that are password protected.

This data set will also be de-identified, by removing the demographic data, and saved on the QUT network drives, which are automatically backed-up nightly in two physical locations. The records of data analysis undertaken and the de-identified data will be stored on github as well as on QUT network drives.

@@

The analysis scripts and the de-identified data will be available for secondary use by other researchers to reproduce these results and also use in future studies. Participants will be asked to provide consent for this usage in the consent form.

@@

A document on the web app is attached.

>>>>

Factor combinations to be examined by each participant amount to 6 (2x3) lineup displays. A participant cannot see the same data for both plot types. Four simulated sets of data will be generated for each treatment. This will generate 24 lineups (12 will be geographic maps, and 12 will be hexagon tile maps). Participants will evaluate 12 lineups, 6 of each plot type. 
Appendix A shows the experimental design visually.

@@

##	Recruitment strategies, timeframe

Recruitment will take place via advertising on the Figure-Eight site.
On this site the available tasks like this survey, are advertised to participants by the addition to the list of tasks available to contributors on the platform. Participants must select our task from the list to read the description of the task, and what is expected of them to be able to participate.

Removed:
[Recruitment will take place via advertising on the Figure-Eight dashboards.
The Figure-Eight platform will advertise this as a "job" to potential participants, they will not be under any obligation to complete this task. We keep advertising until enough participants have completed the survey, then remove the advertisements and close the survey.]

@@

##	Participant withdrawal
Participants can withdraw from the experiment at any time by quitting the survey, they with also be withdrawn if they do not provide a choice for more than 50% of the displays.

@@

Test questions are the most important quality control mechanism in the Figure-Eight Platform. Participants will complete them to practice and prepare for the experiment questions. in our analysis stage, to use the responses from a participant they will need to reach an accuracy threshold of at least one correct choice.

Removed:
[Test questions are the most important quality control mechanism in the Figure-Eight Platform. Participants will complete them to practice and prepare for the experiment questions, they will need to reach an accuracy threshold we will set, to be allowed to participate in answering the experiment survey questions.

The Figure Eight procedures are followed: *Contributors that fail quiz mode are not paid and are permanently disqualified from working on your job. Participants will need to achieve a pre-determined accuracy threshold to be allowed to answer the questions in our survey.*]

@@

Data will be collected through survey distribution via a web application.
Information and illustrative displays regarding the survey have been included in the word document 01_O_ETH_Survey_Data_Collection_Tool_V1_20191023.
Participants will access the application via a link from the Figure-Eight task site.
This will lead them to the web page that currently hosts the pilot study: https://ebsmonash.shinyapps.io/dingo/

@@

```{r echo=FALSE, eval=FALSE, fig.height=5, fig.width=6, out.width="70%"}
# Code to compute the power for given sample size and alternative p
library(tidyverse)
library(nullabor)
library(viridis)
library(knitr)
n <- 20:50
pA <- seq(1/12, 1/3, 0.002)
pow <- sample_size(n, 12, pA)
#g <- expand.grid(n, pA)
#pow <- tibble(n=g[,1], k=as.integer(ceiling(g[,1]/12)), pA=g[,2]) 
#pow <- pow %>% 
#  mutate(prob = pbinom(k-1, n, pA, lower.tail=FALSE)) 
sample_fig <- ggplot(pow, aes(x=n, y=pA, fill=prob, group=pA)) + 
  geom_tile() +
  scale_fill_viridis_c("power") +
  geom_hline(yintercept=0.1666667, colour="black") +
  ylab("detect rate (pA)") + xlab("sample size (n)") +
  theme_bw()
ggsave(filename = "figures/sample_size.jpg", plot = sample_fig, width = 10, height = 10, units = "in")

pow %>% filter(near(pA, 0.167, 0.001)) %>% filter(prob>0.7) %>% top_n(n=10) %>% kable(digits=3)
```

```{r sample_size_fig, out.width = "90%", out.height="90%"}
library(knitr)
library(tidyverse)
knitr::include_graphics("figures/sample_size.jpg")
```

```{r sample_size_table, fig.cap="Chance of detection for sample sizes of 48 and 49."}
tibble(n = c(48, 49),
  k = c(7, 7),
  pA = c(0.167, 0.167),
  prob = c(0.713, 0.734)) %>% kable(digits=3)
```

@@

Research expectations today require that publications use reproducible research protocols, requiring data to be available. The external survey data will be made available via QUT’s Research Data Finder, on github, and, if required by the publisher, once the results are published also available on the journal article web site.

@@

Datasets produced from the project and that will be available for public access on github will be de-identified and only contain the responses to survey, and information regarding the participant demographics: age range, education level, gender, and whether they have lived in Australia.
 If there are participants with unique sets of demographics, their data will be combined with larger categories to minimise the risks of reidentification.

Responses to the survey will be stored securely using a google sheet owned by Stephanie’s Monash University google account drive, accessible only to the researchers during the study through their password protected google drive accounts. This will contain the demographic information, at the coarse resolution specified in the data collection.

Removed:
[Datasets produced from the project and that will not be available for public access will contain original
demographic information about the participants and only the anonymous Figure-Eight contributor ID.
Datasets produced from the project and that will be available for public access on github will be de-identified
and only contain the responses to survey. If there are participants with unique sets of demographics, their
data will be combined with larger categories to minimise the risks of reidentification.]

@@

Removed: 
[Responses to the survey will be stored securely using a private google drive accessible only to the researchers
during the study through their password protected google drive accounts. This will contain the demographic
information, at the coarse resolution specified in the data collection.]

@@

To ensure records are retained during and following the study the responses will be stored on the google sheet immediately following the submission by participants. At the conclusion of the survey, the data will remain in the password protected google drive account.

Immediately following the submission by participants. At the conclusion of the survey, the data will remain in the password protected google drive account and on github
During the analysis phase the researchers Stephanie and Dianne will make copies of the survey responses to their work devices (laptop computers), that are password protected.

@@

Research expectations today require that publications use reproducible research protocols, requiring data to be available. The external survey data will be made available will be made available via QUT’s Research Data Finder, on github, and, if required by the publisher, once the results are published also available on the journal article web site.

The analysis scripts and the data will be available for secondary use by other researchers to reproduce these results and also use in future studies. Participants will be asked to provide consent for this usage in the consent form.

@@

## Appendix A {#sec:appendix_design}

The experimental design plan shows the distribution of the null data sets to each type of trend model that will be tested. 
It also states the location of the trend model in the lineup of 12 plots in a three row, four column grid.

```{r design, fig.width = 10, fig.height=8}
knitr::include_graphics("figures/skobakian_design.pdf")
```
